\documentclass[oneside]{article}

\usepackage[
  paperheight = 9in,
  paperwidth  = 6in,
  textheight  = 7.5in,
  textwidth   = 5in,
]{geometry}

\usepackage{
  amsmath,
  amssymb,
  amsthm,
  anyfontsize, % arbitrary font size
  bm,          % boldsymbol
  bookmark,    % replaces (and loads) hyperref
  fixltx2e,    % textsubscript
  mathrsfs,    % mathscr
  mathtools,   % allows prescript
  microtype,   % improved looks
  tikz-cd,     % better commutative diagrams
  thmtools,    % custom end to example environment
  7370,        % custom style file
}
\usetikzlibrary{matrix,decorations.pathreplacing}

\usepackage[
  hyperref = true,      % links to online documents
  backend  = bibtex,    % use bibtex instead of biber
  sorting  = nyt,       % sorts by (name, year, title)
  style    = alphabetic % citations look like [Har77]
]{biblatex}
\addbibresource{7370.bib}

\hypersetup{
  colorlinks = true,
  linkcolor  = blue,
  urlcolor   = cyan
}

\pagestyle{myheadings}
\markright{\hfill Ranks of elliptic curves\hfill}


\title{Ranks of elliptic curves}
\author{Daniel Miller and Ravi Ramakrishna}
\date{Fall 2014}


\begin{document}

\maketitle
\tableofcontents





\input{1_introduction}
\input{2_background}

\section{Main stuff}

Let $k$ be a number field, $C_{/k}$ a smooth proper curve. Let $g$ be the 
genus of $C$. If $g=0$, either $C(k)=\varnothing$, or $C\simeq \dP^1_{/k}$. 
If $g>1$, then Faltings gave an ineffective proof that $\# C(k)<\infty$. 
If $g=1$, then either $C(k)=\varnothing$, or, if $C(k)\ne\varnothing$, then 
$C(k)$ is a finitely generated abelian group. Recall that the Birch and 
Swinnerton-Dyer conjecture is that $\rank(E(k))=\order_{s=1} L(E,s)$. The 
function $L(E,s)$ satisfies a functional equation $L(E,s) \sim \pm L(E,2-s)$. 
We expect $(-1)^{\rank(E)}=\pm$ accordingly. This is the ``sign conjecture.'' 
Kolyvagin and others have proved that if $L(E,s)$ has a pole of order at most 
$1$ at $1$, then BSD holds for $E$. A folklore question is: is 
$\sup_{E/\dQ} \rank(E)<\infty$?

Any $E_{/\dQ}$ can be written as $y^2=x^3+A x+B$ where $A,B\in \dZ$, and 
$p^6\mid B\Rightarrow p^4\nmid A$. The \emph{height} of $E$ is 
$H(E)=\max(4|A|^3,27 B^2)$. This is asymptotically degree-six in the roots 
of $x^3+A x+B$. We could have ordered $E$ by (Galois-theoretic) conductor. The 
reason we use height is that the number of elliptic curves with height 
$\leqslant X$ is $C X^{5/6}+(\text{error term})$ for some constant $C$. On the 
other hand, we have no idea how many elliptic curves there are with bounded 
conductor. So even though height isn't as conceptually elegant, we'll order 
elliptic curves by it. 

Given $E_{/\dQ}$ with algebraic rank $3$, we could numerically show that the 
analytic rank is $\leqslant 3$, then use sign considerations to get t
$r_\mathrm{an}\in \{1,3\}$. If $r_\mathrm{an}=1$, we know that 
$r_\mathrm{alg}=1$, a contradiction, hence $r_\mathrm{an}=3$. Thus, for 
(specific) elliptic curves with algebraic rank $3$, it is possible to prove 
BSD. At the present, no one knows how to do this for a family of elliptic 
curves. From the Kummer exact sequence $0\to E[p] \to E\xrightarrow p E \to 0$, 
one gets a short exact sequence
\[\begin{tikzcd}
  0 \ar[r] 
    & E(\dQ)/p \ar[r] 
    & \h^1(\dQ,E[p]) \ar[r] 
    & \h^1(\dQ,E)[p] 
    & 0 .
\end{tikzcd}\]
One has a similar short exact sequence for each place $v$. This yields a 
(now-familiar) commutative diagram:
\[\begin{tikzcd}
  0 \ar[r] 
    & E(\dQ)/p \ar[r] \ar[d] 
    & \h^1(\dQ,E[p]) \ar[r] \ar[d] 
    & \h^1(\dQ,E)[p] \ar[d] \ar[r] 
    & 0 \\
  0 \ar[r] 
    & \prod E(\dQ_v)/p \ar[r] 
    & \prod \h^1(\dQ_v,E[p]) \ar[r] 
    & \prod \h^1(\dQ_v,E)[p] \ar[r] 
    & 0.
\end{tikzcd}\]
Even better, we can write 
\[
  \selmer_p(E)  \ker\left(\h^1(\dQ,E[p])\to \h^1(\dA,E)[p]\right) ,
\]
where $\dA$ is the ring of adeles over $\dQ$ and ``$\h^1$'' denotes \'etale 
cohomology. 

We'd like to give a geometric interpretation of $\h^1(\dQ,E)$ and 
$\h^1(\dQ,E[p])$. First, note that 
$\dim (E(\dQ)/p)=\rank(E)+\delta$, where $\delta=\dim E[p](\dQ)\leqslant 2$. 
With respect to height, $\delta>0$ on a density zero set, so we may as well 
assume $\delta=0$. Recall there is a short exact sequence 
\[\begin{tikzcd}
  0 \ar[r] 
    & E(\dQ)/p \ar[r] 
    & \selmer_p(E) \ar[r] 
    & \sha[p] \ar[r] 
    & 0 .
\end{tikzcd}\]
Separating $\selmer_p(E)$ into its contributions from $E(\dQ)/p$ and 
$\sha[p]$ is very difficult. One conjectures that 
$\sha=\ker(\h^1(\dQ,E) \to \h^1(\dA,E))$ is finite for all $E$. This conjecture 
implies that $\sha[p]=0$ for almost all $p$. Cassels has constructed a perfect 
pairing $\sha\times \sha\to \dQ/\dZ$. So if $\#\sha<\infty$, then $\#\sha$ 
should be a perfect square. The ``$\sha$-part'' of BSD can be computed in 
some examples, and it always turns out to be a perfect square. 

[\ldots recap of strategy for Bhargava's proof\ldots]





\newpage
\printbibliography[heading=bibintoc]
\end{document}
